---
title: "Black Friday to New Year’s 2021 in Washington’s CCRS Era"
subtitle: "A First Look at Sales Behavior at Year-End Holiday Season"
date: 2022-01-31
categories: [CCRS, Cannabis, Sales, Holiday]
format: html
editor: 
  markdown: 
    wrap: 72
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}

library(tidyverse)
library(arrow)
library(DBI)
library(duckdb)
library(fs)
library(lubridate)
library(scales)

base_dir <- "C:/Users/theob/OneDrive/Documents/parquet_cannabis_master/arrow"
duckdb_path <- file.path(dirname(base_dir), "duckdb_master.duckdb")

con <- dbConnect(duckdb::duckdb(), dbdir = duckdb_path, read_only = FALSE)

register_master_parquet <- function(con, base_dir) {
  master_files <- list.files(base_dir, pattern = "\\.parquet$", full.names = TRUE)

  table_names <- tools::file_path_sans_ext(basename(master_files)) %>%
    str_replace_all("([a-z])([A-Z])", "\\1_\\2") %>%
    str_replace_all("[^[:alnum:]_]", "_") %>%
    str_to_lower()

  walk2(table_names, master_files, ~ {
    table_name <- paste0(.x, "_master")
    if (table_name %in% duckdb_list_arrow(con)) {
      duckdb_unregister_arrow(con, table_name)
      message("🧹 Unregistered master: ", table_name)
    }

    dataset <- arrow::open_dataset(.y)
    duckdb_register_arrow(con, table_name, dataset)
    message("📦 Registered master: ", table_name)
  })
}

# Run it
register_master_parquet(con, base_dir)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
create_combined_view <- function(con, base_name) {
  tables <- duckdb_list_arrow(con)
  relevant_tables <- tables[str_detect(tables, paste0("^", base_name, "_(master|arrow_.*2025)"))]

  if (length(relevant_tables) == 0) {
    message("⚠️ No tables found for ", base_name)
    return(invisible(NULL))
  }

  # Get Arrow schema columns for each table
  column_lists <- map(relevant_tables, ~ {
  cols <- names(dbGetQuery(con, glue::glue("SELECT * FROM \"{.x}\" LIMIT 0")))
  tibble(table = .x, column = cols)
}) %>% bind_rows()

  # Keep only columns common to all tables
  common_columns <- column_lists %>%
    group_by(column) %>%
    summarise(n_tables = n_distinct(table), .groups = "drop") %>%
    filter(n_tables == length(relevant_tables)) %>%
    pull(column)

  if (length(common_columns) == 0) {
    message("❌ No common columns found across tables for ", base_name)
    return(invisible(NULL))
  }

  # (Optional) Log any dropped columns
  all_cols <- column_lists %>% distinct(column)
  dropped_cols <- setdiff(all_cols$column, common_columns)
  if (length(dropped_cols)) {
    message("⚠️ Dropping columns not in all tables for ", base_name, ": ", paste(dropped_cols, collapse = ", "))
  }

  # Build UNION ALL
  union_parts <- map_chr(relevant_tables, ~ {
    cols <- paste(common_columns, collapse = ", ")
    glue::glue("SELECT {cols} FROM \"{.x}\"")  # quote in case of special chars
  })

  union_sql <- glue::glue("
    CREATE OR REPLACE VIEW {base_name}_all AS
    {paste(union_parts, collapse = '\nUNION ALL\n')}
  ")

  dbExecute(con, union_sql)
  message("✅ Created view: ", base_name, "_all with ", length(relevant_tables), " parts.")
}

views_to_create <- c("sales_detail", "sale_header", "inventory", "product", "licensee")

walk(views_to_create, ~ create_combined_view(con, .x))


```

```{r, echo=FALSE, message=FALSE, warning=FALSE}

sales_blackfriday_xmas_nye_2021 <- dbGetQuery(con, "
  SELECT
    sd.SaleDetailId,
    sh.LicenseeId,
    lic.DBA AS BusinessName,
    sh.SaleDate,
    CAST(sd.Quantity AS DOUBLE) AS Quantity,
    CAST(sd.UnitPrice AS DOUBLE) AS UnitPrice,
    CAST(sd.SalesTax AS DOUBLE) AS SalesTax,
    CAST(sd.OtherTax AS DOUBLE) AS OtherTax,
    CAST(sd.Discount AS DOUBLE) AS Discount,
    
    prod.Name AS ProductName,
    prod.Description AS ProductDescription,
    prod.InventoryType,
    prod.UnitWeightGrams,
    
    (CAST(sd.Quantity AS DOUBLE) * CAST(sd.UnitPrice AS DOUBLE)
     + CAST(sd.SalesTax AS DOUBLE)
     + CAST(sd.OtherTax AS DOUBLE)) AS Revenue

  FROM sales_detail_all sd
  LEFT JOIN sale_header_all sh ON sd.SaleHeaderId = sh.SaleHeaderId
  LEFT JOIN licensee_all lic    ON sh.LicenseeId = lic.LicenseeId
  LEFT JOIN inventory_all inv   ON sd.InventoryId = inv.InventoryId
  LEFT JOIN product_all prod    ON inv.ProductId = prod.ProductId

  WHERE sh.SaleDate BETWEEN '2021-11-24' AND '2021-12-31'
    AND CAST(sd.UnitPrice AS DOUBLE) < 2
")

```

When Washington’s **Cannabis Central Reporting System (CCRS)** went live in December 2021, the transition marked the first statewide attempt to standardize traceability after BioTrack’s phase-out. To understand how licensees were performing during those pivotal early weeks, we analyzed nearly **12 thousand transactions** covering the **Black Friday–Christmas–New Year’s 2021** period.

The dataset captures point-of-sale activity reported by a handful of early CCRS adopters—primarily the Lucid retail group and Green Lady Lacey—and offers a snapshot of how products, pricing, and discounting behaved across inventory types covering nearly **12 thousand transactions** covering the **Black Friday–Christmas–New Year’s 2021** period.

Top Product Types by Revenue
```{r, echo=FALSE, message=FALSE, warning=FALSE}


sales_blackfriday_xmas_nye_2021 %>%
  group_by(InventoryType) %>%
  summarise(
    TotalRevenue = sum(Revenue, na.rm = TRUE),
    UnitsSold = sum(Quantity, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(desc(TotalRevenue)) %>%
  ggplot(aes(x = reorder(InventoryType, TotalRevenue), y = TotalRevenue)) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  scale_y_continuous(labels = dollar_format()) +
  labs(
    title = "💰 Revenue by Product Type (Black Friday - NYE 2021)",
    x = "Product Type",
    y = "Total Revenue"
  ) +
  theme_minimal()


```

Revenue vs Units Sold by Type See if some product types sell lots of units but earn less (or vice versa):

```{r, echo=FALSE, message=FALSE, warning=FALSE}

sales_blackfriday_xmas_nye_2021 %>%
  group_by(InventoryType) %>%
  summarise(
    TotalRevenue = sum(Revenue, na.rm = TRUE),
    TotalUnits = sum(Quantity, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  ggplot(aes(x = TotalUnits, y = TotalRevenue, label = InventoryType)) +
  geom_point(color = "darkgreen", size = 4) +
  geom_text(nudge_y = 5, check_overlap = TRUE) +
  scale_y_continuous(labels = dollar_format()) +
  labs(
    title = "📦 Units Sold vs Revenue by Product Type",
    x = "Units Sold",
    y = "Revenue"
  ) +
  theme_minimal()


```

Daily Sales Trend by Product Type

```{r, echo=FALSE, message=FALSE, warning=FALSE}


sales_blackfriday_xmas_nye_2021 %>%
  mutate(SaleDate = as.Date(SaleDate)) %>%
  group_by(SaleDate, InventoryType) %>%
  summarise(DailyRevenue = sum(Revenue, na.rm = TRUE), .groups = "drop") %>%
  ggplot(aes(x = SaleDate, y = DailyRevenue, color = InventoryType)) +
  geom_line(size = 1) +
  scale_y_continuous(labels = dollar_format()) +
  labs(
    title = "📈 Daily Revenue Trend by Product Type",
    x = "Date",
    y = "Revenue"
  ) +
  theme_minimal()


```

Group by BusinessName and InventoryType This answers questions like: Who sold the most usable cannabis? Which stores had the most revenue from pre-rolls?

```{r, echo=FALSE, message=FALSE, warning=FALSE}

sales_blackfriday_xmas_nye_2021 %>%
  filter(Revenue > 0) %>%
  group_by(BusinessName, InventoryType) %>%
  summarise(
    TotalRevenue = sum(Revenue, na.rm = TRUE),
    TotalUnits = sum(Quantity, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(desc(TotalRevenue)) %>%
  ggplot(aes(x = reorder(BusinessName, TotalRevenue), y = TotalRevenue, fill = InventoryType)) +
  geom_col() +
  coord_flip() +
  scale_y_continuous(labels = dollar_format()) +
  labs(
    title = "💰 Revenue by Product Type and Business",
    x = "Retailer",
    y = "Revenue"
  ) +
  theme_minimal()

```

Normalize Revenue by UnitWeightGrams This gives you revenue per gram — useful for comparing product efficiency, price compression, or margin trends.

```{r, echo=FALSE, message=FALSE, warning=FALSE}

sales_blackfriday_xmas_nye_2021 %>%
  filter(Revenue > 0) %>%
  mutate(UnitWeightGrams = as.numeric(UnitWeightGrams)) %>%
  filter(!is.na(UnitWeightGrams), UnitWeightGrams > 0) %>%
  mutate(RevenuePerGram = Revenue / UnitWeightGrams) %>%
  group_by(InventoryType) %>%
  summarise(
    AvgRevenuePerGram = mean(RevenuePerGram, na.rm = TRUE),
    MedianRevenuePerGram = median(RevenuePerGram, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  ggplot(aes(x = reorder(InventoryType, AvgRevenuePerGram), y = AvgRevenuePerGram)) +
  geom_col(fill = "darkorange") +
  coord_flip() +
  scale_y_continuous(labels = dollar_format()) +
  labs(
    title = "⚖️ Avg Revenue per Gram by Product Type",
    x = "Product Type",
    y = "Revenue per Gram ($)"
  ) +
  theme_minimal()

```

Analyze Discount Behavior by Product Type This identifies which products get discounted most — helps you spot pricing pressure, clearance items, or strategic promotions.

```{r, echo=FALSE, message=FALSE, warning=FALSE}

sales_blackfriday_xmas_nye_2021 %>%
  filter(Revenue > 0) %>%
  group_by(InventoryType) %>%
  summarise(
    AvgDiscount = mean(Discount, na.rm = TRUE),
    MedianDiscount = median(Discount, na.rm = TRUE),
    PercentDiscounted = mean(Discount > 0) * 100,
    .groups = "drop"
  ) %>%
  ggplot(aes(x = reorder(InventoryType, AvgDiscount), y = AvgDiscount)) +
  geom_col(fill = "purple") +
  coord_flip() +
  labs(
    title = "💸 Average Discount by Product Type",
    x = "Product Type",
    y = "Avg Discount ($)"
  ) +
  theme_minimal()


```

Goal 2: Compute Total Revenue per Licensee This answers: Who made the most money during Black Friday → New Year’s Eve 2021?

```{r, message=FALSE, warning=FALSE}


# Clean and prepare the dataset
sales_licensee_summary <- sales_blackfriday_xmas_nye_2021 %>%
  filter(Revenue > 0) %>%
  group_by(LicenseeId, BusinessName) %>%
  summarise(
    TotalRevenue = sum(Revenue, na.rm = TRUE),
    Transactions = n(),
    .groups = "drop"
  ) %>%
  arrange(desc(TotalRevenue))

# View the number of unique licensees
n_licensees <- n_distinct(sales_licensee_summary$LicenseeId)
cat("✅ Number of unique licensees:", n_licensees, "\n")

library(kableExtra)
kable(sales_licensee_summary)

```

<br>


