---
title: "Black Friday to New Year’s 2022 in Washington’s CCRS Era"
subtitle: "Another Look at Sales Behavior at Year-End Holiday Season"
date: "2023-01-29"
categories: [CCRS, Cannabis, Discount, Sales, Holiday]
format: html
fig-height: 4.5
fig-width: 6
toc: true
editor:
  markdown:
    wrap: 72
---
**The Bottom Line Up Front**  
From 5th House Farms’ half-billion-dollar December to the enduring dominance of inhalable concentrates, Washington’s cannabis industry closed 2022 on a powerful note.
The state’s data tell a simple story: the holidays belong to heavy hitters.

For analysts, regulators, and retailers alike, this year-end surge provides a roadmap for what to expect as the market continues to formalize — where data, timing, and product mix will define who wins the next year-end "green holiday season." 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, error = FALSE)

library(DBI)
library(duckdb)
library(tidyverse)
library(plotly)
library(lubridate)
library(scales)
library(kableExtra)

# Connect to DuckDB
source("../R_scripts/connect_duckdb_readonly.R")

start_date <- "2022-11-25"
end_date <- "2023-01-01"

holiday_sales <- dbGetQuery(con, glue::glue("
SELECT
  sh.LicenseeId,
  li.DBA AS BusinessName,
  CAST(sh.SaleDate AS DATE) AS SaleDate,
  p.InventoryType,
  CAST(sd.UnitPrice AS DOUBLE) AS UnitPrice,
  CAST(sd.Discount AS DOUBLE) AS Discount,
  CAST(sd.Quantity AS DOUBLE) AS Quantity,
  CAST(p.UnitWeightGrams AS DOUBLE) AS UnitWeightGrams
FROM SalesDetail_all sd
JOIN SaleHeader_all sh ON sd.SaleHeaderId = sh.SaleHeaderId
LEFT JOIN Licensee_all li ON li.LicenseeId = sh.LicenseeId
LEFT JOIN Inventory_all inv ON sd.InventoryId = inv.InventoryId
LEFT JOIN Product_all p ON inv.ProductId = p.ProductId
WHERE DATE(sh.SaleDate) BETWEEN '{start_date}' AND '{end_date}'
  AND CAST(sd.UnitPrice AS DOUBLE) < 200
"))
holiday_sales <- holiday_sales %>%
mutate(
revenue = UnitPrice * Quantity,
revenue_per_gram = ifelse(UnitWeightGrams > 0, revenue / (Quantity * UnitWeightGrams), NA_real_),
net_revenue = revenue - Discount
)

```
**What It Means for 2023**
**The 2022 holiday data underscore several structural trends that would carry into 2023 and beyond:**  

+ High Concentrate Dependency: The market is top-heavy — over 60% of total value tied to inhalable extracts.

+ Flat Discounting: Despite economic pressure, retailers avoided margin erosion in premium segments.

+ Data Coverage Expansion: More licensees submitted CCRS-compliant data than ever before, improving transparency and market analysis.

Together, these indicators reflect a maturing, efficiency-driven cannabis economy — one increasingly resembling conventional consumer packaged goods (CPG) industries in its seasonal planning, price targeting, and brand hierarchy.

<br>
**Topline Summary**

| Metric                            | Value                                     |
| :-------------------------------- | :---------------------------------------- |
| **Period**                        | Nov 25, 2022 – Jan 1, 2023                |
| **Total Revenue**                 | ≈ $925 million                            |
| **Top Product Category**          | Concentrates for Inhalation ($577 M, 62%) |
| **Number of Reporting Licensees** | ~60                                       |
| **Peak Sales Day**                | December 22, 2022                         |
| **Peak Category Share**           | Concentrates: 65% of total sales          |

<br>
**Overview**

Between Black Friday (Nov 25) and New Year’s Day (Jan 1), Washington’s cannabis market recorded one of its most intense and diverse retail periods to date.
Data from CCRS show more than $925 million in reported transactions across dozens of product types, spanning from traditional usable cannabis to high-end inhalable concentrates and niche tinctures.

The holiday window highlighted both the breadth of Washington’s cannabis supply chain and its evolving retail dynamics — where price, product form, and potency all shaped consumer behavior heading into 2023.

CCRS cannabis sales during the 2022 holiday season, specifically from Black Friday (Nov 25, 2022) through New Year’s Day (Jan 1, 2023).

**This ledger post covers:**

1. Revenue by product type

2. Revenue vs. units sold by product type

3. Daily sales trends by product type

4. Business and inventory type grouping

5. Revenue per gram (normalized by weight)

6. Discount behavior by product type

7. Total revenue per licensee

**We can now answer:**

Who sold the most usable cannabis or pre-rolls?

Which product type had the most discounting behavior?

How efficient was each product in terms of revenue per gram?

<br>
**Revenue by Product Type**
```{r}

holiday_sales %>%
group_by(InventoryType) %>%
summarise(total_revenue = sum(revenue, na.rm = TRUE)) %>%
arrange(desc(total_revenue)) %>%
mutate(total_revenue = dollar(total_revenue)) %>%
kable(caption = "Total Revenue by Product Type") %>%
kable_styling(full_width = FALSE)


```
<br>

**Revenue vs. Units Sold by Product Type**
```{r, fig.width=10}

library(scales)
library(ggrepel)

holiday_sales %>%
  group_by(InventoryType) %>%
  summarise(
    total_revenue = sum(revenue, na.rm = TRUE),
    total_units = sum(Quantity, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    total_revenue = pmin(total_revenue, 60000000)
  ) %>%
  filter(!is.na(InventoryType), total_units > 0, total_revenue > 0) %>%
  ggplot(aes(x = total_units, y = total_revenue, label = InventoryType)) +
  geom_point(color = "forestgreen", size = 3, alpha = 0.8) +
  ggrepel::geom_text_repel(size = 3.5, max.overlaps = 20) +
  scale_x_log10(labels = comma_format()) +
  scale_y_log10(labels = dollar_format()) +
  labs(
    title = "Revenue vs. Units Sold by Product Type (Log Scale)",
    x = "Units Sold (log10)",
    y = "Total Revenue (log10)"
  ) +
  theme_minimal()

```
<br>

**Daily Sales Trends by Product Type**
```{r}

daily_trends <- holiday_sales %>%
group_by(SaleDate, InventoryType) %>%
summarise(total_revenue = sum(revenue, na.rm = TRUE), .groups = "drop") %>%
  mutate(total_revenue = pmin(total_revenue, 60000000)) %>%
  filter(!is.na(InventoryType))

plot_ly(daily_trends,
x = ~SaleDate,
y = ~total_revenue,
color = ~InventoryType,
type = 'scatter',
mode = 'lines+markers') %>%
layout(
title = "Daily Revenue Trends by Product Type",
xaxis = list(title = "Date"),
yaxis = list(title = "Revenue (USD)")
)


```
<br>

**What Consumers Bought**

Concentrates for Inhalation absolutely dominated, generating $577 million, or nearly two-thirds of total holiday sales.  Traditional usable cannabis (flower) followed distantly at $76 million, while edibles (solid and liquid combined) made up under 1% of overall sales.

The dominance of extract-based products underscores how established consumers — especially those purchasing in bulk or for gifting — have shifted toward higher-potency forms.

<br>

**Revenue by Business + Product Type**
```{r, fig.width=10}

holiday_sales %>%
group_by(BusinessName, InventoryType) %>%
summarise(total_revenue = sum(revenue, na.rm = TRUE), .groups = "drop") %>%
arrange(desc(total_revenue)) %>%
slice_head(n = 20) %>%
ggplot(aes(x = reorder(BusinessName, total_revenue), y = total_revenue, fill = InventoryType)) +
geom_col() +
coord_flip() +
scale_y_continuous(labels = dollar) +
labs(
title = "Top 20 Licensees by Revenue and Product Type",
x = "Business Name",
y = "Total Revenue"
) +
theme_minimal()

```
<br>

**Revenue per Gram by Product Type**
```{r}

holiday_sales %>%
filter(!is.na(revenue_per_gram)) %>%
group_by(InventoryType) %>%
summarise(
avg_revenue_per_gram = mean(revenue_per_gram, na.rm = TRUE)
) %>%
arrange(desc(avg_revenue_per_gram)) %>%
mutate(avg_revenue_per_gram = dollar(avg_revenue_per_gram)) %>%
dplyr::rename("Revenue Per Gram (Mean)" = avg_revenue_per_gram, "Inventory Type" = InventoryType) %>%
kable(caption = "Average Revenue per Gram by Product Type") %>%
kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))

```
<br>

**Price and Value: How Much Was a Gram Worth?**

During the holiday window, Transdermals and CO₂ concentrates fetched the highest prices per gram, reflecting niche medical and artisanal segments.  The price stratification shows a clear tiering of Washington’s cannabis economy — from everyday flower under $4/gram to boutique topicals and extracts exceeding $80/gram.

**This divergence reveals a key insight:**  
Retailers protected margin on high-demand inhalables while discounting slower-moving, niche categories like tinctures and topicals to clear shelf space before year-end.

<br>

**Discount Behavior by Product Type**
```{r}

holiday_sales %>%
group_by(InventoryType) %>%
summarise(
avg_discount_pct = sum(Discount, na.rm = TRUE) / sum(revenue, na.rm = TRUE) * 100
) %>%
arrange(desc(avg_discount_pct)) %>%
mutate(avg_discount_pct = percent(avg_discount_pct / 100, accuracy = 0.1)) %>%
dplyr::rename("Discount Percent (Mean)" = avg_discount_pct, "Inventory Type" = InventoryType) %>%
kable(caption = "Average Discount % by Product Type") %>%
kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))

```
<br>

**Discount Behavior: Who Marked Down for the Holidays**

Discounting varied widely across product categories.
While tinctures and non-solvent concentrates saw double-digit markdowns, inhalable concentrates — the main revenue driver — showed virtually no discounting.

<br>

**Total Revenue per Licensee**
```{r}


holiday_sales %>%
group_by(BusinessName) %>%
summarise(total_revenue = sum(revenue, na.rm = TRUE)) %>%
arrange(desc(total_revenue)) %>%
mutate(total_revenue = dollar(total_revenue)) %>%
dplyr::rename("Total Revenue" = total_revenue, "Business Name" = BusinessName) %>%
kable(caption = "Total Revenue per Licensee – Holiday Season 2022") %>%
kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))

```
<br>

**Who Drove the Market: Top Holiday Licensees**  
Holiday sales were highly concentrated — literally and figuratively.
The top 10 licensees accounted for nearly 85% of total sales, led by 5th House Farms, Seattle’s Private Reserve, and Mt. Baker Homegrown.

These producers and processors form Washington’s “holiday backbone” — scaling up concentrate and extract output to meet December demand surges.

**The Calendar of the Holiday Rush**  
**Analysis of daily sales data shows two major waves:**

+ Black Friday Spike (Nov 25–29): modest volume dominated by flower and edibles.

+ Mid-December Explosion (Dec 8–22): multiple high-value concentrate surges, with record-breaking single-day totals exceeding $60 million (Dec 8).

+ Sales fell sharply on Christmas Day and spiked again from aDec 26–30, ending with elevated New Year’s Eve totals.

Holiday demand and inventory management strategy must mirror mainstream retail trends; early promotional engagement, mid-December inventory runs, and late-month re-stocking by retailers before the post-holiday season and typical January slowdown.

