---
title: "Encapsulated Flow: Mapping Transport Flows in Washington’s Capsule Market"
subtitle: "Production Cycles and Distribution Patterns (2021–2025)"
description: |
  This analysis examines the movement of cannabis capsule products within 
  Washington’s regulated supply chain. Capsule-specific trends reveal where losses,
  bottlenecks, and high-volume transfer hubs emerge among licensees, offering
  actionable insight into capsule operations and regulatory oversight.
date: "2025-12-28"
image: true
categories: [Capsule, Supply Chain, Transportation, 4Years]
project:
  type: website
format: 
  html: 
    toc: true
    code-fold: false
    theme: flatly
editor:
  markdown:
    wrap: 72
---
```{r, echo=FALSE}
library(htmltools)

tags$style(HTML("
/* Make DT tables fit content area better */
.dataTables_wrapper {
font-size: 0.9rem;
}

/* Reduce row height */
table.dataTable tbody td {
padding: 6px 8px;
vertical-align: middle;
}

/* Improve header contrast */
table.dataTable thead th {
background-color: #f8f9fa;
border-bottom: 2px solid #dee2e6;
font-weight: 600;
white-space: nowrap;
}

/* Keep IDs readable */
td.dt-nowrap {
white-space: nowrap;
font-family: monospace;
}

/* Sticky filter row spacing */
.dataTables_filter {
margin-bottom: 0.5rem;
}

/* Caption styling */
caption {
font-size: 0.85rem;
color: #6c757d;
}
"))
```

## The Bottom Line Up Front  
The Market Is Concentrated but Stable - a small number of origin licensees anchor capsule distribution statewide. This concentration introduces dependency risk, but also suggests operational maturity. Capsule distribution favors localized production-to-retail loops, likely driven by predictable demand, shelf stability, limited wholesale arbitrage, and higher regulatory scrutiny relative to inhaler products.

## Overview  
Capsules occupy a unique position in Washington’s cannabis market. They are non-inhalable, dosage-controlled, medically adjacent, and operationally conservative compared to flower or concentrates. Yet, despite their relatively modest market share, capsule products generate a rich and revealing transportation footprint inside the state’s seed-to-sale system.

Using CCRS transport manifests, inventory records, and adjustment logs, this analysis traces how capsule products move across Washington, from origin licensees through distribution corridors to retail destinations. The result is a detailed look at logistics behavior, regional hubs, and supply chain concentration for one of the market’s most compliance-sensitive product types.

```{r setup, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}

library(tidyverse)
library(DBI)
library(dplyr)
library(lubridate)
library(plotly)
library(arrow)
library(duckdb)
library(fs)
library(glue)
library(kableExtra)
library(scales)
library(sf)
library(tigris)
library(ggrepel)
library(osrm)
library(DT)

options(tigris_use_cache = TRUE)

if (exists("con")) 
  try(DBI::dbDisconnect(con), silent = TRUE)

source("../R_scripts/connect_duckdb_readonly.R")

## ---------------------------------------------------------
## Licensee Table
## ---------------------------------------------------------

licensee_ref <- dbGetQuery(
  con,
  "
  SELECT
    LicenseeId,
    Name  AS BusinessName,
    DBA   AS DoingBusinessAs
  FROM licensee_clean
  WHERE LOWER(IsDeleted) = 'false'
  "
)

## ---------------------------------------------------------
## Capsule Inventory Investigation Pipeline
## ---------------------------------------------------------

## 1. Pull capsule products --------------------------------
capsule_products <- dbGetQuery(
  con,
  "
  SELECT
    ProductId,
    LicenseeId,
    InventoryType,
    Name AS ProductName,
    UnitWeightGrams,
    IsDeleted
  FROM product_clean
  WHERE LOWER(InventoryType) = 'capsule'
   AND LOWER(IsDeleted) = 'false'
  "
)

## 2. Inventory tied to capsule products -------------------
capsule_inventory <- dbGetQuery(
  con,
  "
  SELECT
    InventoryId,
    LicenseeId,
    ProductId,
    AreaId,
    InitialQuantity,
    QuantityOnHand,
    TotalCost,
    IsMedical,
    CreatedDate,
    IsDeleted
  FROM inventory_clean
  WHERE LOWER(IsDeleted) = 'false'
  "
)

capsule_inventory <- capsule_inventory %>%
  left_join(
    licensee_ref,
    by = "LicenseeId"
  )

capsule_inventory <- 
  capsule_inventory %>%
  inner_join(capsule_products, by = "ProductId", suffix = c(".inv", ".prod"))

## 3. Area context -----------------------------------------
areas <- dbGetQuery(
  con,
  "
  SELECT
    AreaId,
    LicenseeId,
    Name AS AreaName,
    IsQuarantine
  FROM areas_all
  WHERE LOWER(IsDeleted) = 'false'
  "
)

capsule_inventory <- capsule_inventory %>%
  left_join(areas, by = c(
  "AreaId" = "AreaId",
  "LicenseeId.inv" = "LicenseeId"))

## 4. Inventory adjustments --------------------------------
inventory_adjustments <- dbGetQuery(
  con,
  "
  SELECT
    InventoryId,
    InventoryAdjustmentReason,
    Quantity,
    AdjustmentDate
  FROM InventoryAdjustment_all
  WHERE LOWER(IsDeleted) = 'false'
  "
)

capsule_adjustments <- capsule_inventory %>%
  left_join(inventory_adjustments, by = "InventoryId")

# ## 5. Lab results ------------------------------------------
# lab_results <- dbGetQuery(
#   con,
#   "
#   SELECT
#     InventoryId,
#     TestName,
#     TestValue,
#     TestDate,
#     LabTestStatus
#   FROM LabResult_all
#    WHERE LOWER(IsDeleted) = 'false'
#   "
# ) 
# 
# capsule_labs <- capsule_inventory %>%
#   left_join(lab_results, by = "InventoryId")

## 6. Sales activity ---------------------------------------
sales_detail <- dbGetQuery(
  con,
  "
  SELECT
    SaleDetailId,
    SaleHeaderId,
    InventoryId,
    Quantity,
    UnitPrice,
    Discount,
    SalesTax
  FROM SalesDetail_all
  WHERE LOWER(IsDeleted) = 'false'
  "
) 

capsule_sales <- capsule_inventory %>%
  left_join(sales_detail, by = "InventoryId") %>%
  mutate(
    GrossRevenue = as.numeric(Quantity, na.rm = TRUE) * as.numeric(UnitPrice, na.rm = TRUE),
    NetRevenue = as.numeric(GrossRevenue, na.rm = TRUE) - as.numeric(Discount, na.rm = TRUE)
  )

## 7. Transportation (capsule manifests) -------------------
transported_capsules <- dbGetQuery(
  con,
  "
  SELECT
    ExternalManifestIdentifier,
    InventoryExternalIdentifier,
    InventoryType,
    Quantity,
    UOM,
    CreatedDate
  FROM TransportedItems_all
  WHERE LOWER(InventoryType) = 'capsule'
    AND LOWER(IsDeleted) = 'false'
  "
)

## Inventory health summary -----------------------------
capsule_inventory_summary <- capsule_inventory %>%
  group_by(LicenseeId.inv, ProductName) %>%
  summarise(
    InventoryRecords = n(),
    TotalInitialQty = sum(as.numeric(InitialQuantity), na.rm = TRUE),
    TotalOnHand = sum(as.numeric(QuantityOnHand), na.rm = TRUE),
    .groups = "drop"
  )

## Adjustment reason breakdown --------------------------
capsule_adjustment_summary <- capsule_adjustments %>%
  group_by(InventoryAdjustmentReason) %>%
  summarise(
    TotalAdjustedQty = sum(as.numeric(Quantity), na.rm = TRUE),
    Adjustments = n(),
    .groups = "drop"
  )

manifest_headers <- dbGetQuery(
  con,
  "
  SELECT
    ExternalManifestIdentifier,
    OriginLicenseNumber,
    OriginLicenseName,
    DestinationLicenseNumber,
    DestinationLicenseName,
    DepartureDateTime,
    ArrivalDateTime
  FROM ManifestHeader_all
  WHERE LOWER(IsDeleted) = 'false'
    AND LOWER(OrderCancelled) = 'false'
  "
) 

licensees <- dbGetQuery(
  con,
  "
  SELECT
    LicenseeId,
    LicenseNumber,
    Name AS LicenseeName,
    City,
    County,
    State,
    ZipCode
  FROM licensee_all
  WHERE LOWER(IsDeleted) = 'false'
  "
)

capsule_movements <- transported_capsules %>%
  inner_join(
    manifest_headers,
    by = "ExternalManifestIdentifier"
  ) %>%
  mutate(
    DepartureDate = as_date(DepartureDateTime),
    ArrivalDate   = as_date(ArrivalDateTime)
  )

```
```{r, include=FALSE}

capsule_flows <- capsule_movements %>%
  left_join(
    licensees,
    by = c("OriginLicenseNumber" = "LicenseNumber"),
    relationship = "many-to-many"
  ) %>%
  rename_with(~ paste0("Origin_", .),
              c(LicenseeId, LicenseeName, City, County, State, ZipCode)) %>%
  left_join(
    licensees,
    by = c("DestinationLicenseNumber" = "LicenseNumber"),
    relationship = "many-to-many"
  ) %>%
  rename_with(~ paste0("Destination_", .),
              c(LicenseeId, LicenseeName, City, County, State, ZipCode))

capsule_flows <- capsule_flows %>% 
  distinct(ExternalManifestIdentifier, InventoryExternalIdentifier, .keep_all = TRUE)

# Licensee-Level Capsule Flow Aggregation
capsule_flow_summary <- capsule_flows %>% 
    group_by(
    Origin_LicenseeId,
    Origin_LicenseeName,
    Origin_County,
    Destination_LicenseeId,
    Destination_LicenseeName,
    Destination_County
  ) %>%
  summarise(
    Shipments = n_distinct(ExternalManifestIdentifier),
    TotalUnits = sum(as.numeric(Quantity), na.rm = TRUE),
    AvgUnitsPerShipment = mean(as.numeric(Quantity), na.rm = TRUE),
    FirstShipment = min(DepartureDate, na.rm = TRUE),
    LastShipment  = max(DepartureDate, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  filter(Origin_LicenseeId != Destination_LicenseeId)

# County-to-County Flow Map Table
capsule_county_flows <- capsule_flow_summary %>%
  group_by(
    Origin_County,
    Destination_County
  ) %>%
  summarise(
    Shipments = sum(Shipments),
    TotalUnits = sum(TotalUnits),
    .groups = "drop"
  ) %>%
  filter(!is.na(Origin_County), !is.na(Destination_County))

# High-Volume Capsule Hubs
capsule_hubs <- capsule_flow_summary %>%
  group_by(Origin_LicenseeId, Origin_LicenseeName) %>%
  summarise(
    OutboundShipments = sum(Shipments),
    OutboundUnits = sum(TotalUnits),
    TradingPartners = n_distinct(Destination_LicenseeId),
    .groups = "drop"
  ) %>%
  arrange(desc(OutboundUnits))

```
## Licensee-Level Capsule Flow Summary
*Who ships capsules to whom, and how much?*

::: {.column-page}
```{r capsule-flow-table, message=FALSE, warning=FALSE, echo=FALSE, include=TRUE}

datatable(
  capsule_flow_summary %>%
    mutate(
      FirstShipment = as.character(FirstShipment),
      LastShipment  = as.character(LastShipment)
    ),
  extensions = c("Buttons", "FixedHeader"),
  options = list(
    pageLength = 15,
    autoWidth = TRUE,
    fixedHeader = TRUE,
    dom = "Bfrtip",
    buttons = c("copy", "csv", "excel"),
    order = list(list(6, "desc")), # TotalUnits
    columnDefs = list(
      list(className = "dt-right", targets = c(6, 7))
    )
  ),
  rownames = FALSE,
  filter = "top",
  caption = htmltools::tags$caption(
    style = "caption-side: bottom; text-align: left;",
    "Licensee-level capsule trading relationships. ",
    htmltools::em("Sorted and filterable.")
  )
) %>%
  formatRound(
    columns = c("Shipments", "TotalUnits", "AvgUnitsPerShipment"),
    digits = 0,
    mark = ","
  )

```
:::  

## Who Ships Capsules at Scale
```{r, include=TRUE, echo=FALSE}

top_hubs_n <- 15

capsule_hubs %>%
  slice_max(OutboundUnits, n = top_hubs_n) %>%
  mutate(Origin_LicenseeName = fct_reorder(Origin_LicenseeName, OutboundUnits)) %>%
  ggplot(aes(x = OutboundUnits, y = Origin_LicenseeName)) +
  geom_col(fill = "#1f78b4") +
  scale_x_continuous(labels = comma) +
  labs(
    title = "Top Capsule Shipping Hubs",
    subtitle = paste0("Top ", top_hubs_n, " licensees by outbound units"),
    x = "Outbound Units",
    y = NULL
  ) +
  theme_minimal(base_size = 12)

```

## High-Volume Capsule Hubs  

::: {.column-page}
```{r capsule-hubs-table, message=FALSE, warning=FALSE, echo=FALSE, include=TRUE}

datatable(
  capsule_hubs,
  extensions = c("Buttons", "FixedHeader"),
  options = list(
    pageLength = 15,
    autoWidth = TRUE,
    fixedHeader = TRUE,
    dom = "Bfrtip",
    buttons = c("copy", "csv", "excel"),
    order = list(list(3, "desc")) # OutboundUnits
  ),
  rownames = FALSE,
  filter = "top",
  caption = htmltools::tags$caption(
    style = "caption-side: bottom; text-align: left;",
    "Top capsule distribution hubs ranked by outbound volume."
  )
) %>%
  formatRound(
    columns = c("OutboundShipments", "OutboundUnits"),
    digits = 0,
    mark = ","
  )

```
:::

## Licensee Hubs: Few Origins, Many Destinations  
Outbound shipment data reveals a hub-and-spoke structure:

+ Just 32 origin licensees function as major capsule hubs  
+ Some hubs ship to nearly 200 trading partners
+ Others operate as vertically aligned loops, shipping primarily to themselves or affiliates  

The largest hubs are not necessarily the largest producers by volume, but rather those with:

+ stable SKUs,  
+ repeatable packaging formats,  
+ and consistent compliance performance.

This structure reduces transport complexity but increases **network dependency** on a small set of operators.

**What These Tables Tell Us About the Capsule Market** - capsules are a logistics-conservative product  
Capsule supply chains favor:

+ short distances,

+ known partners,

+ predictable demand,

+ and compliance stability.

They behave more like **regulated consumer health products** than traditional cannabis SKUs.
```{r, message=FALSE, warning=FALSE, include=FALSE}
# WA county geometry
# Reproject counties to planar CRS before centroid calculation
wa_counties_sf <- tigris::counties(state = "WA", cb = TRUE, year = 2022) %>%
  st_as_sf() %>%
  st_transform(3857) %>%   # Web Mercator (safe for centroids)
  transmute(
    County = str_to_upper(NAME),
    GEOID,
    geometry
  )

wa_centroids <- wa_counties_sf %>%
  mutate(centroid = st_point_on_surface(geometry)) %>%
  mutate(
    cx = st_coordinates(centroid)[, 1],
    cy = st_coordinates(centroid)[, 2]
  ) %>%
  st_drop_geometry()

```

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}

county_out <- capsule_county_flows %>%
  group_by(Origin_County) %>%
  summarise(OutboundUnits = sum(TotalUnits, na.rm = TRUE), .groups = "drop") %>%
  rename(County = Origin_County)

county_in <- capsule_county_flows %>%
  group_by(Destination_County) %>%
  summarise(InboundUnits = sum(TotalUnits, na.rm = TRUE), .groups = "drop") %>%
  rename(County = Destination_County)

county_net <- wa_counties_sf %>%
  left_join(county_out, by = "County") %>%
  left_join(county_in, by = "County") %>%
  mutate(
    OutboundUnits = replace_na(OutboundUnits, 0),
    InboundUnits  = replace_na(InboundUnits, 0),
    NetFlowUnits  = OutboundUnits - InboundUnits
  )

p_net_interactive <- ggplot(county_net) +
  geom_sf(aes(
    fill = NetFlowUnits,
    text = paste0(
      "County: ", str_to_title(County),
      "<br>Outbound Units: ", comma(OutboundUnits),
      "<br>Inbound Units: ", comma(InboundUnits),
      "<br>Net Units: ", comma(NetFlowUnits)
    )
  ), color = "white", linewidth = 0.2) +
  scale_fill_gradient2(low = "#2b8cbe", mid = "grey95", high = "#e34a33", midpoint = 0) +
  labs(title = "Capsule Net Flow by County (Interactive)", fill = "Net Units") +
  theme_minimal(base_size = 12) +
  theme(panel.grid = element_blank(), axis.text = element_blank(), axis.title = element_blank())

plotly::ggplotly(p_net_interactive, tooltip = "text")

```

```{r, warning=FALSE, message=FALSE, include=FALSE}

# Pick top N lanes to keep the map readable
top_n_lanes <- 75

flow_lines <- capsule_county_flows %>%
  mutate(
    Origin_County = str_to_upper(Origin_County),
    Destination_County = str_to_upper(Destination_County)
  ) %>%
  inner_join(wa_centroids, by = c("Origin_County" = "County")) %>%
  rename(ox = cx, oy = cy) %>%
  inner_join(wa_centroids, by = c("Destination_County" = "County")) %>%
  rename(dx = cx, dy = cy) %>%
  arrange(desc(TotalUnits)) %>%
  slice_head(n = top_n_lanes) %>%
  mutate(
    lane_label = paste0(str_to_title(Origin_County), " → ", str_to_title(Destination_County)),
    w = scales::rescale(log1p(TotalUnits), to = c(0.2, 2.5)) # stroke width
  )

flow_lines <- flow_lines %>%
  mutate(
    tooltip = paste0(
      "Route: ", lane_label,
      "<br>Shipments: ", scales::comma(Shipments),
      "<br>Total Units: ", scales::comma(TotalUnits)
    ),
    color_val = log1p(TotalUnits)   # better visual spread
  )

p_routes <- ggplot() +
  geom_sf(
    data = wa_counties_sf,
    fill = "grey97",
    color = "grey55",
    linewidth = 0.6
  ) +
  geom_segment(
    data = flow_lines,
    aes(
      x = ox, y = oy,
      xend = dx, yend = dy,
      linewidth = w,
      color = color_val,
      key = tooltip
    ),
    arrow = arrow(type = "closed", length = unit(2, "mm")), 
    alpha = 0.75,
    lineend = "round"
  ) +
  scale_linewidth_identity() +
  scale_color_viridis_c(
    option = "C",
    labels = scales::comma,
    name = "Capsule Units\n(log scale)"
  ) +
  labs(
    title = "Top Capsule Trading Lanes (County-to-County)",
    subtitle = paste0("Top ", top_n_lanes, " lanes by total units"),
    caption = "Line width and color scaled by shipment volume"
  ) +
  theme_void(base_size = 12) +
  theme(
    legend.position = "right",
    plot.title = element_text(face = "bold")
  )

plotly::ggplotly(p_routes, tooltip = "key")

```
## "Lane Drilldown" tables (for compliance traceability)  

::: {.column-page}
```{r, echo=FALSE, include=TRUE}

# A drilldown table: top manifest lines by quantity, includes External IDs
capsule_top_manifest_lines <- capsule_flows %>%
  transmute(
    ExternalManifestIdentifier,
    InventoryExternalIdentifier,
    Origin_LicenseeName,
    Destination_LicenseeName,
    Origin_County,
    Destination_County,
    Quantity = as.numeric(Quantity),
    UOM,
    DepartureDate
  ) %>%
  arrange(desc(Quantity)) %>%
  slice_head(n = 50)

datatable(
  capsule_top_manifest_lines %>%
    mutate(across(where(is.character), ~ replace_na(.x, ""))),
  extensions = c("Buttons", "FixedHeader"),
  options = list(
    pageLength = 15,
    autoWidth = TRUE,
    fixedHeader = TRUE,
    dom = "Bfrtip",
    buttons = c("copy", "csv", "excel"),
    order = list(list(6, "desc")), 
    columnDefs = list(
      list(className = "dt-right", targets = c(6)) # numeric alignment
    )
  ),
  rownames = FALSE,
  filter = "top",
  caption = htmltools::tags$caption(
    style = "caption-side: bottom; text-align: left;",
    "Top Capsule Manifest Lines (with External Identifiers)"
  )
) %>%
  formatRound(
    columns = c("Quantity"),
    digits = 0
  ) %>%
  formatStyle(
  "Quantity",
  background = DT::styleColorBar(
    capsule_top_manifest_lines$Quantity,
    "#2c7fb8"
  ),
  backgroundSize = "98% 88%",
  backgroundRepeat = "no-repeat",
  backgroundPosition = "center"
) %>%
formatStyle(
  c("ExternalManifestIdentifier", "InventoryExternalIdentifier"),
  fontWeight = "bold",
  color = "#1f78b4"
)

```
:::

## Mapping Capsule Transport Routes  

::: {.column-page}
```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}

wa_centroids_ll <- wa_centroids %>%
  st_as_sf(coords = c("cx", "cy"), crs = 3857) %>%
  st_transform(4326)

route_pairs <- flow_lines %>%
  select(
    Origin_County,
    Destination_County,
    tooltip,
    color_val,
    w
  ) %>%
  left_join(wa_centroids_ll, by = c("Origin_County" = "County")) %>%
  rename(orig_pt = geometry) %>%
  left_join(wa_centroids_ll, by = c("Destination_County" = "County")) %>%
  rename(dest_pt = geometry)

route_pairs <- route_pairs %>%
  mutate(route_id = row_number())

routes_list <- purrr::pmap(
  list(route_pairs$route_id,
       route_pairs$orig_pt,
       route_pairs$dest_pt),
  function(id, orig, dest) {
    tryCatch(
      {
        r <- osrmRoute(
          src = st_as_sf(st_sfc(orig, crs = 4326)),
          dst = st_as_sf(st_sfc(dest, crs = 4326)),
          overview = "full"
        )
        r$route_id <- id
        r
      },
      error = function(e) NULL
    )
  }
)

routes_sf <- routes_list %>%
  purrr::compact() %>%
  bind_rows() %>%
  left_join(
    route_pairs %>%
      select(route_id, w, color_val, tooltip),
    by = "route_id"
  )

p <- ggplot() +
  geom_sf(
    data = wa_counties_sf,
    fill = "grey97",
    color = "grey55",
    linewidth = 0.6
  ) +
  geom_sf(
    data = routes_sf,
    aes(
      linewidth = w,
      color = color_val,
      key = tooltip
    ),
    alpha = 0.8
  ) +
  scale_linewidth_identity() +
  scale_color_viridis_c(option = "C",
                        name = "Capsule Units\n(log\u2081\u2080)") +
  theme_void()

```
:::  

Transport Data Reveals Strategy, Not Just Movement - even without pricing or margin data, transport flows expose:

+ vertical integration strategies,
+ geographic retail footprints,
+ and regional demand consistency.

## Interactive Capsule Transport Map (with tooltips) 

::: {.column-page}
```{r, message=FALSE, warning=FALSE, echo=FALSE, include=TRUE}

p_interactive <- plotly::ggplotly(p, tooltip = "key")

p_interactive
```
:::

## County-to-County Capsule Flows
```{r capsule-county-table, message=FALSE, warning=FALSE, echo=FALSE, include=TRUE}

datatable(
  capsule_county_flows,
  extensions = c("Buttons", "FixedHeader"),
  options = list(
    pageLength = 10,
    autoWidth = TRUE,
    fixedHeader = TRUE,
    dom = "Bfrtip",
    buttons = c("copy", "csv", "excel"),
    order = list(list(3, "desc")) # TotalUnits
  ),
  rownames = FALSE,
  filter = "top",
  caption = htmltools::tags$caption(
    style = "caption-side: bottom; text-align: left;",
    "Aggregated capsule movement between Washington counties."
  )
) %>%
  formatRound(
    columns = c("Shipments", "TotalUnits"),
    digits = 0,
    mark = ","
  )

```

## County Corridors: The Spine of Capsule Logistics

While capsule movements are widespread, a small number of county corridors account for a disproportionate share of volume.

High-volume corridors include:

+ Whatcom → Whatcom

+ Whatcom → King

+ Skagit → Snohomish

+ Pacific → King

+ Thurston → Pierce

These routes combine short distance, established operators, and reliable retail endpoints, reinforcing the idea that capsule logistics prioritize operational certainty over expansion.

## Why This Matters  
For operators, capsule logistics show where predictable revenue lives.  
For regulators, they highlight low-risk compliance corridors.  
For analysts, they offer a clean contrast to the volatility seen in flower and concentrate markets.

Most importantly, this analysis demonstrates the power of open CCRS data: even routine manifest records can reveal how the cannabis economy actually functions — product by product, route by route.

This is one of a continuing series of transport and supply-chain analyses published by The Evergreen Canna Ledger. All findings are derived from anonymized, aggregated CCRS data and are intended for transparency, research, and public understanding.