---
title: "Washington’s 4/20 Evolution: From Niche Retailers to $128 Million Month — April 2025 Sets a New Benchmark"
subtitle: "Four Years April Cannabis Summary: 2022–2025"
date: "2025-05-01"
categories: [Cannabis, Discount, Revenue, Holiday, 4Years]
format: html
toc: true
editor:
  markdown:
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, error = FALSE)

library(DBI)
library(duckdb)
library(tidyverse)
library(plotly)
library(lubridate)
library(scales)
library(kableExtra)

# Connect to DuckDB
source("../R_scripts/connect_duckdb_readonly.R")

april_monthly <- dbGetQuery(con, "
WITH base AS (
SELECT
sh.LicenseeId,
DATE_TRUNC('month', CAST(sh.SaleDate AS DATE)) AS Month,
CAST(sd.UnitPrice AS DOUBLE) AS UnitPrice,
CAST(sd.Quantity AS DOUBLE) AS Quantity,
CAST(sd.Discount AS DOUBLE) AS Discount
FROM SalesDetail_all sd
JOIN SaleHeader_all sh ON sd.SaleHeaderId = sh.SaleHeaderId
WHERE strftime(CAST(sh.SaleDate AS DATE), '%Y-%m') IN
('2022-04', '2023-04', '2024-04', '2025-04')
AND CAST(sd.UnitPrice AS DOUBLE) < 200
)
SELECT
strftime(Month, '%Y-%m') AS YearMonth,
COUNT(DISTINCT LicenseeId) AS licensee_reporting,
ROUND(SUM(UnitPrice * Quantity), 2) AS total_revenue,
ROUND(SUM(Discount) / NULLIF(SUM(UnitPrice * Quantity), 0) * 100, 2) AS avg_discount_pct
FROM base
GROUP BY YearMonth
ORDER BY YearMonth
")
```
**The Bottom Line Up Front**  
From $1.5 M in 2022 to $127.9 M in 2025, April’s cannabis revenue surge reflects not just consumer enthusiasm, but the structural maturity of Washington’s legal cannabis market.
What began as a handful of retailers reporting sales during the 4/20 holiday is now a comprehensive economic snapshot of a fully functioning, highly competitive industry.

As CCRS participation and analytical visibility continue improving, April 2025 may be remembered as the moment Washington’s cannabis data went from fragmented experiment to economic benchmark.


**Market Interpretation**
What does this mean for the broader cannabis economy?

**Scale & Data Coverage:**  
The 2025 data mark the first true baseline year for statewide analytics — virtually the entire legal market is now visible through CCRS.

**Price Compression:**  
The jump in discounting percentages indicates growing competition and narrower margins, typical of a stabilizing retail sector.

**Predictable Seasonality:**  
The 4/20 period has become a reliable economic indicator, offering early insight into annual sales trajectories, similar to how December holiday spending predicts retail performance in other industries.

**Policy & Planning:**  
For regulators and industry groups, the rapid increase in reporting licensees underscores the success of data transparency mandates and POS integration efforts over the past two years.



**Overview**  
Every April, Washington’s cannabis retailers brace for the industry’s biggest day — 4/20, the unofficial consumer holiday celebrating cannabis culture.  From 2022 through 2025, the state’s CCRS sales data paint a vivid picture of how far the market has come: from a small cluster of early-reporting licensees to a full-scale, competitive retail ecosystem generating nine-figure monthly revenues.


**Monthly Comparison Table**  

**Total Revenue by April**

| Year | Licensees Reporting | Total Revenue | Avg. Discount % |
| :--- | ------------------: | ------------: | --------------: |
| 2022-04 |                  14 |        $1.5 M |           9.7 % |
| 2023-04 |                  16 |        $5.2 M |           0.5 % |
| 2024-04 |                  62 |        $8.5 M |           5.1 % |
| 2025-04 |           **1,045** |  **$127.9 M** |      **12.4 %** |

<br>

In just three years, April sales expanded over 80-fold, a reflection of:

+ The growing number of CCRS-compliant retailers,

+ Improved reporting automation across point-of-sale systems, and

+ Increased consumer participation around the 4/20 promotional period.

While 2022–2023 were still in the data-onboarding phase, 2024 and especially 2025 mark the first truly complete statewide picture of 4/20 performance.

```{r}
 
april_weekly <- dbGetQuery(con, "
WITH base AS (
SELECT
sh.LicenseeId,
li.DBA AS BusinessName,
strftime(CAST(sh.SaleDate AS DATE), '%Y-%m') AS YearMonth,
DATE_TRUNC('week', CAST(sh.SaleDate AS DATE)) AS WeekStart,
CAST(sd.UnitPrice AS DOUBLE) AS UnitPrice,
CAST(sd.Quantity AS DOUBLE) AS Quantity,
CAST(sd.Discount AS DOUBLE) AS Discount
FROM SalesDetail_all sd
JOIN SaleHeader_all sh ON sd.SaleHeaderId = sh.SaleHeaderId
LEFT JOIN Licensee_all li ON li.LicenseeId = sh.LicenseeId
WHERE strftime(CAST(sh.SaleDate AS DATE), '%Y-%m') IN
('2022-04', '2023-04', '2024-04', '2025-04')
AND CAST(sd.UnitPrice AS DOUBLE) < 200
)
SELECT
YearMonth,
WeekStart,
COUNT(DISTINCT LicenseeId) AS licensee_reporting,
ROUND(SUM(UnitPrice * Quantity), 2) AS total_revenue,
ROUND(SUM(Discount) / NULLIF(SUM(UnitPrice * Quantity), 0) * 100, 2) AS avg_discount_pct
FROM base
GROUP BY YearMonth, WeekStart
ORDER BY YearMonth, WeekStart
")

```
<br>
**Weekly Summaries within Each April**

**Weekly Plot: Revenue per Week (Aprils Only)**
```{r}

plot_ly(
april_weekly,
x = ~WeekStart,
y = ~total_revenue,
color = ~YearMonth,
type = "scatter",
mode = "lines+markers"
) %>%
layout(
title = "Weekly Revenue Trends for April (2022–2025)",
xaxis = list(title = "Week"),
yaxis = list(title = "Total Revenue (USD)")
)

```

**Weekly Plot: Discount % per Week**
```{r}

plot_ly(
april_weekly,
x = ~WeekStart,
y = ~avg_discount_pct,
color = ~YearMonth,
type = "scatter",
mode = "lines+markers"
) %>%
layout(
title = "Weekly Discount % for April (2022–2025)",
xaxis = list(title = "Week"),
yaxis = list(title = "Average Discount %")
)

```
<br>  

**The Weekly Story Behind 4/20**

| Year |          Peak Week | Revenue (USD) | Avg. Discount % |
| :--- | -----------------: | ------------: | --------------: |
| 2022 |        Apr 18 – 24 |       $5.47 M |           5.1 % |
| 2023 |        Apr 17 – 23 |       $12.4 M |           1.3 % |
| 2024 |        Apr 15 – 21 |       $19.6 M |           6.2 % |
| 2025 | **Mar 31 – Apr 6** |    **$310 M** |      **20.8 %** |

<br>

Each year, Washington’s 4/20 sales activity has shifted slightly earlier in April — a by-product of longer promotional calendars and pre-holiday discounting.

+ In 2022, the 4/20 surge was modest but clear: a 5 M-plus week among fewer than 10 retailers.

+ By 2023, that surge doubled to more than 12 M in sales, still among just 15 licensees.

+ 2024 expanded both participation (≈40 retailers) and total April sales, setting the stage for scale.

+ And in 2025, Washington reached a new era: over 800 active retail reporters generated $310 M during the first week of April alone, followed by a sustained monthlong sales wave above $100 M.

<br>
**Discounting Behavior Around 4/20**  
Discounting has become the key competitive lever during April.
In 2022–2023, most retailers kept discounts conservative (5–10 %), targeting margin stability.
By 2025, however, discounts averaged 12 % across April, peaking near 21 % in the first week, as retailers fought to capture early-season shoppers.

The trend suggests a more mature price war dynamic:

Larger operators leveraging volume promotions,

Smaller boutiques differentiating through niche products rather than deep discounts, and

Statewide consumer sensitivity to holiday deals rivaling mainstream retail events like Black Friday.

<br>
**Weekly Summary**  
```{r}

april_weekly %>%
mutate(
avg_discount_pct = percent(avg_discount_pct / 100, accuracy = 0.1),
total_revenue = dollar(total_revenue)
) %>%
arrange(YearMonth, WeekStart) %>%
kable(
caption = "Weekly Summary for April 2022–2025",
col.names = c("Month", "Week", "Licensees", "Total Revenue", "Avg. Discount %")
) %>%
kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))

```

<br>

