---
title: "The Doomed Product Detector: Which Cannabis SKUs Are About to Disappear?"
subtitle: "Identifying Products at Risk of Discontinuation"
image: false
date: "2025-11-30"
categories: [CCRS, Product, Lifecycle]
project:
  type: website
format: 
  html: 
    toc: true
    code-fold: false
    theme: flatly
editor:
  markdown:
    wrap: 72
---

**The Bottom Line Up Front**  
A more data-driven future for Washington’s competitive, saturated, and constantly evolving cannabis economy, identified doomed products before they hit reset bins.  

**This way stakeholders can:**

+ streamline purchasing

+ reduce losses

+ improve category management

+ make strategic decisions based on real statewide data

**Overview**  
Every January, WA retailers refresh their shelves — and thousands of slow‑moving SKUs quietly disappear.

This annual “reset cycle” removes products that no longer sell, have aged out of rotation, or have accumulated too much on-hand inventory.

Using 22+ million CCRS product-store movement records of 5.6+ million cannabis products and retail sales velocity metrics between August 04 - November 03, 2025, we modeled which active products show patterns consistent with impending extinction.

We identify “doomed” products by comparing:

+ Sales over the last 4 weeks vs. 13 weeks

+ Product velocity (units sold per store per week)

+ Inventory pressure (quantity on hand vs demand)

+ Market penetration (store count carrying the product)

Products with low velocity, shrinking demand, low penetration, and high inventory form a clear cluster of high‑risk items.

The analysis covers a 91-day window dating backward 4- and 13- weeks from 11-03-2025.

**Table of dates used in this report:**  

| Metric                | Dates Used            |
| --------------------- | --------------------- |
| **Analysis End Date** | Nov 03, 2025          |
| **4-Week Window**     | Oct 05 → Nov 03, 2025 |
| **13-Week Window**    | Aug 04 → Nov 03, 2025 |
| **4-Week Length**     | 28 days               |
| **13-Week Length**    | 91 days               |

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

library(dplyr)
library(readr)
library(kableExtra)
library(here)

# Load preprocessed data
doomed_products <- readRDS("../data/doomed_products.rds")
store_risk <- readRDS("../data/store_risk.rds")
last_sold <- readRDS("../data/last_sold.rds")

```
## Building the “Doom Score”

**A simple, intuitive rule‑of‑thumb score:**

+ Low velocity (bottom quartile)

+ Declining trend (4w < 13w normalized)

+ High inventory pressure (Quantity-On-Hand (QOH): Velocity Ratio)

+ Low market penetration

When all four conditions align, the “doom_flag” is triggered.

These SKUs will likely be delisted before spring resets or stocked until best buy date has passed.

## Doomed Product Table  
Only showing 25 of 4,915 products as example: 
```{r, include=TRUE}

kable(
doomed_products %>%
filter(doom_flag == "Likely Doomed") %>%
select(InventoryType, Name, inv_ratio) %>%
distinct(Name, .keep_all = TRUE) %>%
arrange(desc(inv_ratio)) %>%
rename(Inventory_Ratio = inv_ratio, Inventory_Type = InventoryType) %>%
head(25),
caption = "Top 25 Doomed Products by Inventory Ratio"
)

```

**4,915 products** were found to be likely doomed. 

### Contact us for the complete list. 

## Table of Doom by Inventory Type
```{r, include=TRUE}

#doomed_products %>% select(Name, InventoryType, doom_flag) %>% kable()

kable(doomed_products %>%
  filter(doom_flag != "Healthy") %>%
  distinct(Name, .keep_all = TRUE) %>%
  select(Name, InventoryType, doom_flag) %>%
  count(InventoryType, doom_flag, sort = TRUE) %>%
  rename(Count = n, Inventory_Type = InventoryType, Doom_Flag = doom_flag)
)

```

## Shelf‑Staleness Indicator

Measure time since last sale:
```{r, include=TRUE}

last_sold %>% distinct(Name, .keep_all = TRUE) %>% arrange(desc(Days_Since_Last_Sale)) %>% head(50) %>% kable(caption = "Table of Product Market Staleness") 

```
Short list of products stalling out of market in the last 91 days.

**Some stores have:**

+ 50–70% of initial units still sitting in inventory.

+ Others have sold through most stock, indicating the product is dying market-wide, not store-specific.

+ A few retailers carry only 1–3 doomed SKUs, indicating tight inventory control.


The core insight is most “Doomed” products are not bad products — 

They are orphans.   

**They don’t have:**

+ enough stores

+ enough demand

+ enough velocity

+ enough brand support

+ enough shelf space

Even strong brands produce doomed items if a SKU fails to scale.

## Top 20 Overstocked Products
```{r, include=TRUE}

doomed_products %>%
  filter(doom_flag == "Likely Doomed") %>%
  arrange(desc(inv_ratio)) %>%
  select(Name, inv_ratio, trend_units) %>%
  rename(Inventory_Ratio = inv_ratio, Trend_Units = trend_units) %>%
  distinct(Name) %>%
  head(20) %>%
  kable()

```
**High inventory ratio numbers reflects overstocked products.**

### Retailer Exposure: 

We identified 282 retailers currently carrying measurable quantities of doomed SKUs.

### Store-level Risk: which stores carry the most doomed products?
**The table displays only 20 of the 282 stores as an example**
```{r, include=TRUE}

# Filter doomed products
# View top at-risk stores
store_risk %>%
arrange(desc(doomed_product_count)) %>%
head(20) %>% 
kable(caption = "Top 20 Retailers by Number of Doomed Products")
#kable(store_risk %>% head(20))
# industry benchmarking
#store_risk %>% mean(pct_sold)
#store_risk %>% mean(pct_on_hand)
```

```{r}

rm(doomed_products, store_risk, last_sold)

```

### Contact us for the full report. 

## Methodological Confidence 

**Why trust this model?**

+ Built from millions of CCRS movement lines

+ Uses velocity, penetration, inventory pressure, and trend units

+ Flags products consistently across categories

+ Aligns with observed January reset behavior across retailers

+ Identifies dead inventory down to specific LicenseeIds

Combined with TECL’s open-data commitment, this gives WA industry a transparent, predictable way to understand SKU lifecycle risk. The WA cannabis industry has had no open, systematic way to identify which products are at risk before they vanish.

The Evergreen Canna Ledger’s Doomed Product Detector changes that.

In addition to the above report, TECL can also provide:

2️⃣ Escape Velocity Failure

Products newly introduced in January–March that never break into 5+ stores within the first 60 days.

3️⃣ Price Collapse

Rapid price drops over the last 60–90 days often signal overstock or liquidation.

4️⃣ Retailer Drop‑Off

Track how many stores have stopped ordering a product over the last 8 weeks.

The Evergreen Canna Ledger exists so operators — big and small — can rely on open, transparent, and independently analyzed data instead of rumors or vendor claims.

<br>

{{< include _includes/commento.html >}}
