---
title: "The Price of Pot: Trends in Unit Prices by Product Type"
subtitle: "An analysis of cannabis product inventory levels across Washington state"
date: "2025-11-02"
categories: [CCRS, Cannabis, Price]
project:
  type: website
format: 
  html: 
    toc: true
    code-fold: false
    theme: flatly
editor:
  markdown:
    wrap: 72
---

```{r setup, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}

library(tidyverse)
library(DBI)
library(dplyr)
library(lubridate)
library(plotly)
library(arrow)
library(duckdb)
library(fs)
library(glue)
library(kableExtra)
library(scales)

if (exists("con")) 
    try(DBI::dbDisconnect(con), 
silent = TRUE)

source("C:/Users/theob/OneDrive/Documents/evergreen-canna-ledger/R_scripts/connect_duckdb_readonly.R")



```
## The Bottom Line Up Front  
This four-year pricing dataset reveals a cannabis market that is highly stratified, increasingly efficient, and maturing toward CPG dynamics.
For operators and analysts, the trends highlight where margin pressure is likely to persist—and where premium tiers still justify investment.

**Overview**  
Washington’s cannabis marketplace continues to evolve, shaped by tightening margins, shifting consumer behavior, and seasonal dynamics.
Using nearly four years of transaction-level data (Dec 2021–Nov 2025), we can track how average unit prices have moved across product categories—including flower, edibles, concentrates, tinctures, capsules, and emerging specialty products.

The trends below give operators, analysts, and regulators a clearer view of where price compression is occurring—and where premiums are holding strong.

**Insights**

+ Which categories are trending up or down in price?

+ Are price drops tied to overproduction or seasonal changes?

+ Do some categories show stability compared to others?
```{r, echo=FALSE}
# Query average unit price by month and product category

pricing_trends <- dbGetQuery(con, "
  SELECT
    DATE_TRUNC('month', CAST(sd.CreatedDate AS TIMESTAMP)) AS SaleMonth,
    p.InventoryType,
    AVG(CAST(sd.UnitPrice AS DOUBLE)) AS AvgUnitPrice
  FROM SalesDetail_clean sd
  JOIN Inventory_clean i ON sd.InventoryId = i.InventoryId
  JOIN Product_clean p ON i.ProductId = p.ProductId
  WHERE
    sd.UnitPrice IS NOT NULL AND
    sd.CreatedDate IS NOT NULL AND
    p.InventoryType IS NOT NULL AND
    DATE_TRUNC('month', CAST(sd.CreatedDate AS TIMESTAMP)) 
      BETWEEN DATE '2021-12-01' AND CURRENT_DATE
  GROUP BY SaleMonth, p.InventoryType
  ORDER BY SaleMonth
")

pricing_trends <- pricing_trends %>%
  filter(SaleMonth >= as.Date("2021-12-01") & SaleMonth <= Sys.Date())

```
## Plot Pricing Trends Over Time
```{r, warning=FALSE, echo=FALSE, include=TRUE, fig.width=15, fig.height=25}

plot_ly(
  data = pricing_trends,
  x = ~as.Date(SaleMonth),
  y = ~AvgUnitPrice,
  color = ~InventoryType,
  type = 'scatter',
  mode = 'lines',
  line = list(width = 2)
) %>%
  layout(
    title = "Average Unit Price Over Time by Product Type",
    xaxis = list(title = "Month"),
    yaxis = list(
      title = "Average Unit Price ($)",
      tickformat = "$,.2f"
    ),
    legend = list(orientation = "h", x = 0.1, y = -0.2),
    margin = list(b = 100),
    hovermode = "x unified"
  )

```

## 1. Flower Pricing: Volatile but Stabilizing

**Usable Cannabis (Flower)** — historically the core of Washington’s retail market — shows seasonal fluctuations but a clear long-term downward trend.

- **Late 2021–2022:** Prices ranged **$17–$25**, with early 2022 peaking (e.g., **$25.33 in Jan 2022**).
- **2023:** Stabilized around **$16–$24**, including a spring tightening tied to inventory cycles.
- **2024–2025:** Prices gradually compressed toward **$17–$20**, consistent with increasing competition and fall harvest cycles.

**Flower lots** (bulk/wholesale flower) often show **near-zero values** due to:  

- partial lots  
- returns  
- destruction entries  
- compliance-only adjustments  

These do *not* reflect retail pricing.

**Interpretation:**  
The flower market is mature. Oversupply, aggressive price competition, and consumer trade-down behavior have kept prices flat despite inflationary pressures elsewhere.

---

## 2. Edibles: Remarkably Stable Across the Market

Both solid and liquid edibles show one of the strongest stability profiles in the entire dataset.

### **Solid Edibles**
- Typically range **$16.0–$19.2** over four years.
- Minor monthly fluctuations with no long-term inflation or deflation.

### **Liquid Edibles**
- Generally fall between **$10.5–$15.5**.
- Slight price compression in 2024, with several months dipping **below $11.00**.

**Interpretation:**  
Edibles behave like a classic **consumer packaged goods (CPG)** category — less exposed to grower cycles and more dependent on branding, formulation, dosing, and manufacturing consistency.

---

## 3. Concentrates: Diverging Price Tiers

Washington’s concentrate market clearly splits into distinct segments:

### **1. Premium, Non-Solvent Extracts**
- Often priced **$30–$70** per unit.
- Repeated peaks (e.g., **$69.68 in Mar 2023**, sustained **$53–$57** through 2024–25).
- Driven by strong demand for rosin and solventless SKUs.

### **2. Hydrocarbon & Ethanol Extracts**
- Much more affordable, typically **$6–$20**.
- Occasional deep dips (**$3–$4**) during processor inventory resets.

### **3. CO₂ Extracts**
- Mid-tier pricing, usually **$18–$27**.

### **4. Concentrate for Inhalation (catch-all category)**
- Averages **$14–$28**.
- Moderate inflation in 2023, easing back down in 2024–25.

**Interpretation:**  
Concentrates follow a **dual-market structure**: 

- A **craft premium tier** (solventless, high labor, low yield)  
- A **value tier** (hydrocarbon & CO₂ extracts)  

Demand elasticity is very different between the two segments.

---

## 4. Tinctures and Capsules: High Margin, High Volatility

### **Tinctures**
- Increase steadily over time.
- Late 2022 prices **$14–$23**, rising to **$27–$31** by 2024.
- One of the highest-margin categories.

### **Capsules**
- Typically **$11–$28**.
- Frequent spikes tied to batch availability and production gaps.

**Interpretation:**  
These medical-adjacent categories serve niche but loyal consumers, making them more prone to supply-driven volatility.

---

## 5. Topicals and Transdermals: A Quiet but Consistent Premium Category

### **Topical Ointments**
- Commonly priced **$16–$26**.
- Stable even during broader market price compression.

### **Transdermals**
- Range **$8–$30**, with visible upward drift during 2023–24.

**Interpretation:**  
These categories cater to wellness-oriented consumers and show more resilient price floors than inhalable products.

---

## 6. Emerging, Niche, and Anomalous Product Categories

### **Suppositories**
- Extremely high prices (**$120–$150**).
- Reflect tiny batch sizes and limited competition.

### **Flower Lots & Wet Flower**
Very low averages due to operational entries such as:  

- internal transfers  
- waste  
- partial units  
- destruction manifests  

These are **not** indicators of retail price levels.

### **Plants, Clones, Seeds**
- Near-zero averages are expected; these are typically compliance units, not retail items.

---

## 7. The Big Picture: What Operators Should Take Away

Across **2021–2025**, several consistent themes emerge.

### **1. Price Compression Is Persistent but Uneven**
- Flower and hydrocarbon concentrates face ongoing downward pressure.
- Edibles and topicals remain stable.
- Solventless extracts and tinctures continue to hold premium value.

### **2. Consumers Are Migrating Toward Value**
Recurring dips in 2023–2024 reflect:  

- heavier discounting  
- larger pack sizes  
- consumer trade-down strategies  

### **3. Premiumization Still Exists**
Mostly in:  

- solventless concentrates  
- capsules  
- select tinctures  

### **4. Inventory & Seasonal Cycles Matter**
Annual **Q4 outdoor harvests** correlate with:  

- lower flower prices  
- more concentrate production  
- increased volatility in wholesale categories  

### **5. A Mature, Competitive Market**
Washington resembles:  

- a **commodity market** for flower and inhalables  
- a **CPG market** for edibles and topicals  

This dual identity defines current pricing dynamics.

---

What was provided was a data-driven look at how the price of pot behaves in Washington’s regulated market — and what those patterns mean for cultivators, processors, retailers, and policymakers.

<br>

{{< include _includes/commento.html >}}

