---
title: "Category Evolution 2021–2025: How Washington’s Cannabis Market Grew, Flattened, and Specialized"
subtitle: "From Flower to Formulation in Washington’s Cannabis Categories Evolved from Craft to Commerce"
date: "2025-07-03"
categories: [CCRS, Cannabis, Sales, Product, 4Years]
format: html
fig-height: 4.5
fig-width: 6
toc: true
editor:
  markdown:
    wrap: 72
---

**The Bottom Line Up Front**  
By 2025, Washington’s cannabis sector has matured into a multi-tier retail ecosystem:

+ Commodities at the base,

+ Concentrates as the profit engine, and

+ Edibles as the mass-market frontier.

Where 2021 looked like a boutique craft industry, 2025 looks like a full retail economy — one that behaves more like alcohol, coffee, or CPG snack categories than a novelty substance.

The future now depends less on legalization headlines and more on branding, vertical integration, and consumer segmentation.

**Structural Insights**  
Flower Is Fully Commoditized
Volume up 10× since 2022, average price <$2/gram. Producers survive on efficiency and contracting.

Concentrates Are the Cash Engine
Price stability and margin integrity show brand loyalty and cross-category versatility (vape, dab, cart).

Edibles Are the New Growth Driver
Edibles now account for nearly 15% of statewide revenue, mirroring beverage and snack segmentation.

Niche and Wellness Formats Are Stabilizing
Topicals, tinctures, and transdermals maintain small but steady segments — a predictable, mature structure akin to supplements.

**Overview**  
Over five years, Washington’s legal cannabis market transformed from a small, high-margin boutique sector into a scaled agricultural-retail ecosystem moving billions in volume.

Using CCRS data from 2021–2025, this analysis tracks **category-level performance** — who led, who lagged, and where the state’s cannabis economy is heading.

The numbers tell a dramatic story:

+ Flower volumes exploded, but average prices fell below $2 per gram.

+ Concentrates and edibles gained pricing power, driving premium revenue growth.

+ Infused and packaged mixes emerged as mid-tier staples bridging convenience and potency.

+ Topicals, tinctures, and niche forms maintained small but loyal markets.

<br>
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, error = FALSE)

library(DBI)
library(duckdb)
library(tidyverse)
library(lubridate)
library(scales)
library(kableExtra)
library(plotly)

source("../R_scripts/connect_duckdb_readonly.R")

# Query category-level trends
category_summary <- dbGetQuery(con, "
  WITH sales_base AS (
    SELECT 
      sd.SaleDetailId,
      sh.SaleDate,
      pr.InventoryType,
      CAST(sd.Quantity AS DOUBLE) AS Quantity,
      CAST(sd.UnitPrice AS DOUBLE) AS UnitPrice,
      CAST(sd.Discount AS DOUBLE) AS Discount
    FROM SalesDetail_all sd
    JOIN SaleHeader_all sh ON sd.SaleHeaderId = sh.SaleHeaderId
    JOIN Inventory_all inv ON sd.InventoryId = inv.InventoryId
    JOIN Product_all pr ON inv.ProductId = pr.ProductId
    WHERE (sd.IsDeleted IS NULL OR sd.IsDeleted NOT IN ('TRUE', 'true'))
      AND (sh.IsDeleted IS NULL OR sh.IsDeleted NOT IN ('TRUE', 'true'))
      AND (inv.IsDeleted IS NULL OR inv.IsDeleted NOT IN ('TRUE', 'true'))
      AND (pr.IsDeleted IS NULL OR pr.IsDeleted NOT IN ('TRUE', 'true'))
      AND CAST(sd.UnitPrice AS DOUBLE) < 200
  )

  SELECT 
    EXTRACT(YEAR FROM DATE(SaleDate)) AS Year,
    InventoryType,
    SUM(Quantity) AS UnitsSold,
    ROUND(SUM(Quantity * UnitPrice), 2) AS Revenue,
    ROUND(AVG(UnitPrice), 2) AS AvgUnitPrice
  FROM sales_base
  GROUP BY 1, 2
  ORDER BY Year, Revenue DESC
")

# Format for publication
category_summary %>%
  mutate(
    Revenue = dollar(Revenue),
    AvgUnitPrice = dollar(AvgUnitPrice)
  ) %>%
  rename(
    `Year` = Year,
    `Category` = InventoryType,
    `Units Sold` = UnitsSold,
    `Total Revenue` = Revenue,
    `Avg. Unit Price` = AvgUnitPrice
  ) %>%
  arrange(Category, Year) %>%
  kable("html", caption = "Product Category Trends (2021–2024)") %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))

```

```{r, include=FALSE}

category_summary_clean <- category_summary %>%
  mutate(
    Year = as.integer(Year),
    InventoryType = str_trim(InventoryType),
    Revenue = as.numeric(Revenue),
    UnitsSold = as.numeric(UnitsSold),
    AvgUnitPrice = as.numeric(AvgUnitPrice)
  )

# Identify top categories by total revenue (sum over all years)
top_categories <- category_summary_clean %>%
  group_by(InventoryType) %>%
  summarise(total_revenue = sum(Revenue, na.rm = TRUE)) %>%
  arrange(desc(total_revenue)) %>%
  slice_head(n = 10) %>%
  pull(InventoryType)

# Filter to top categories only for clarity
category_top <- category_summary_clean %>%
  filter(InventoryType %in% top_categories)

```

```{r}

# --- 1. Revenue over time
revenue_plot <- ggplot(category_top, aes(x = Year, y = Revenue, color = InventoryType)) +
  geom_line(size = 1.2) +
  geom_point(size = 2) +
  scale_y_continuous(labels = label_dollar(scale = 1e-6, suffix = "M")) +
  labs(
    title = "Total Revenue by Category Over Time",
    y = "Revenue (Millions)",
    x = "Year",
    color = "Category"
  ) +
  theme_minimal()

revenue_plotly <- ggplotly(revenue_plot)

revenue_plotly

```

<br>
**Edibles: From Niche to Mainstream**  
Edibles — both solid and liquid — surged in 2024–2025.

| Year | Solid Edible Revenue | Liquid Edible Revenue | Avg Unit Price |
| :--- | -------------------: | --------------------: | -------------: |
| 2022 |                $1.9M |                $0.86M |         $12–17 |
| 2023 |                $8.5M |                 $1.9M |         $12–18 |
| 2024 |                $334M |                 $164M |         $10–17 |
| 2025 |            **$436M** |             **$203M** |     **$10–17** |

<br>
**Observation**  
The “snack-ification” of cannabis has arrived.
Solid edibles (chews, baked goods) are now a half-billion-dollar category, while liquid edibles and drinkables are approaching **$200M**.
The steady per-unit pricing shows **consumer normalization** — edibles are no longer a novelty, they’re a staple.

<br>
```{r}
# --- 2. Units sold over time
units_plot <- ggplot(category_top, aes(x = Year, y = UnitsSold, color = InventoryType)) +
  geom_line(size = 1.2) +
  geom_point(size = 2) +
  scale_y_continuous(labels = label_number(scale = 1e-6, suffix = "M")) +
  labs(
    title = "Units Sold by Category Over Time",
    y = "Units Sold (Millions)",
    x = "Year",
    color = "Category"
  ) +
  theme_minimal()

units_plotly <- ggplotly(units_plot)

units_plotly

```

<br>
**Mid-Tier Expansion: Infused & Packaged Mixes**  
Once obscure, Cannabis Mix Infused and Cannabis Mix Packaged now represent nearly $250M combined revenue in 2025.

+ Infused Mix: $167.8M (up 19% YoY)

+ Packaged Mix: $84M (down slightly from 2024’s $106M)

These “bridge categories” combine flower with extracts or pre-packaged convenience — catering to value-driven consumers seeking ready-to-use products.
Average unit prices hover near $9–10, confirming a stable mid-shelf positioning.

**The Premium Edge: High-End Concentrates and Topicals**  

| Product                     | Avg Unit Price (2025) | Revenue |
| :-------------------------- | --------------------: | ------: |
| **Ethanol Concentrate**     |                $22.80 |  $49.8M |
| **Topical Ointment**        |                $22.27 |  $41.3M |
| **Non-Solvent Concentrate** |                $30.50 |  $12.3M |
| **CO₂ Concentrate**         |                $20.60 |  $13.5M |


These small-volume but high-margin categories anchor the upper end of the retail price spectrum, serving medical and wellness-oriented buyers.
Notably, the Non-Solvent Concentrate segment has sustained $30+/unit pricing for three consecutive years, outperforming all other concentrate types on per-unit value.


**Commodities and Byproducts: The Low-End Expansion**  
At the other end of the curve, Flower Lot, Unlotted Flower, and Other Material categories exploded in sheer volume. This confirms a mature supply-chain phase — raw cannabis is now treated like a commodity crop.

| Category           | 2025 Units Sold | Avg Price | Comment                           |
| :----------------- | --------------: | --------: | :-------------------------------- |
| Flower Lot         |            490M |     $1.12 | Bulk agricultural product         |
| Flower Unlotted    |            233M |     $0.41 | Likely processor throughput       |
| Other Material Lot |            205M |     $0.60 | Byproduct sales and waste streams |

The price compression at this end underwrites the margin room for processors and retailers at the top.

<br>
**The Big Two: Flower and Concentrates**  
**Useable Cannabis**

+ 2021: 38k units, $701K revenue, $18.16/unit

+ 2025: 123 million units, $1.57B revenue, $18.59/unit

Despite volume expansion of >3,000x, flower pricing has remained remarkably stable in retail terms — a sign of commoditization offset by branding and consistency.
The data suggest growers are competing on scale, while retailers rely on product differentiation (strain lineage, terpene labeling) to sustain mid-teens per-unit pricing.

**Concentrate for Inhalation**

+ 2021: $36K revenue

+ 2022: $74M

+ 2024: $1.36B

+ 2025: $1.75B

Concentrates remain Washington’s economic backbone, accounting for roughly **40–45% of all cannabis revenue up to May 2025**.

Average price per gram has held around $17–19, even as total volume climbed beyond 130 million units — a rare instance where scale didn’t crush price integrity.
Consumers appear willing to pay for potency and convenience.

**Interpretation:**  
Concentrates are now the oil of the cannabis economy — high-margin, logistics-friendly, and increasingly brand-defined.

<br>
**Emerging Niches and Curiosities**  
While they barely register in total volume, several niche categories are showing steady experimentation and price elasticity:

+ Suppository: $74.51/unit, up 30% YoY

+ Transdermal: $11.54/unit, small but consistent volume

+ Capsules: ~$14/unit, maintaining medical market traction

+ Clones & Seeds: <$3/unit, reflecting agricultural base demand

These reflect the diversification of Washington’s cannabis economy — from high-THC recreational products to wellness, cultivation, and R&D inputs.

<br>
```{r}

# --- 3. Avg. unit price
price_plot <- ggplot(category_top, aes(x = Year, y = AvgUnitPrice, color = InventoryType)) +
  geom_line(size = 1.2) +
  geom_point(size = 2) +
  scale_y_continuous(labels = dollar_format()) +
  labs(
    title = "Avg. Unit Price by Category Over Time",
    y = "Average Unit Price (USD)",
    x = "Year",
    color = "Category"
  ) +
  theme_minimal()

price_plotly <- ggplotly(price_plot)

price_plotly

```
<br>
The pricing landscape now mirrors consumer packaged goods:
commodity base, strong mid-tier competition, and premium niches maintaining identity through quality and consistency.

<br>
**Market Growth at a Glance**

| Year     | Total Revenue | Top Category                   | Avg. Price Range (Top 3 Categories) |
| :------- | ------------: | :----------------------------- | :---------------------------------- |
| **2021** |       ≈ $1.0M | Usable Cannabis                | $15–18 per unit                     |
| **2022** |          $95M | Concentrate for Inhalation     | $9–16                               |
| **2023** |          $55M | Usable Cannabis                | $17–21                              |
| **2024** |         $3.3B | Concentrate for Inhalation     | $10–19                              |
| **Jan-May 2025** |     **$4.2B** | **Concentrate for Inhalation** | **$10–18**                  |

<br>
From 2021’s early-stage artisanal market to 2025’s billion-dollar scale, the data reveal a tenfold jump in total revenue and a massive broadening of licensee participation, especially post-2023 when full CCRS compliance took hold.

<br>


