---
title: "Behind the Brands: A Look at Wholesalers in Washington's Cannabis Industry"
subtitle: "How Washington’s Wholesalers Built a Two-Tier System of Scale and Specialization"
date: "2025-09-21"
categories: [CCRS, Cannabis, Wholesale, Product Diversity, 4Years]
format: html
fig-width: 8
fig-height: 5
toc: true
editor:
  markdown:
    wrap: 72
---

**The Bottom Line Up Front**  
Washington’s wholesale cannabis market is becoming a mirror image of its retail tier:  

+ Broad distributors resemble supermarkets — moving volume with minimal branding.
+ Small “Other” licensees act like boutique suppliers — innovating at the genetic or formulation level.

Together, they form a dual-engine supply chain — one optimized for efficiency, the other for experimentation.
The numbers may be quiet, but they reveal a sophisticated balance between scale, specialization, and brand invisibility — the hidden architecture of Washington’s cannabis economy. 

| Trend                            | Interpretation                                                                                                               |
| :------------------------------- | :--------------------------------------------------------------------------------------------------------------------------- |
| **Flower Dominates Production**  | Even among non-flower specialists, most distributors maintain flower SKUs for margin stability.                              |
| **Clones and Seeds Rising**      | Genetic and starter material categories (plants, clones, seeds) are expanding quietly — often outside consumer view.         |
| **Concentrates Stay High-Value** | Concentrate for Inhalation leads SKU diversity among smaller wholesalers, suggesting sustained product development interest. |
| **Private Label Pipeline**       | The low self-branding rate points to continued demand for private-label or co-packed products.                               |

The wholesale layer is evolving into a modular supply system — one where cultivators, extractors, and distributors collaborate fluidly, rather than vertically integrate.


**Understanding the Dataset**
We analyzed 1,388 wholesale licensees, combining production, product registration, and branding data through mid-2025.

Key indicators included:

+ Total SKUs per licensee

+ Distinct product types (InventoryType)

+ First and last created dates

+ First and last sale dates

+ Branding behavior (SelfBranded, PercentSelfBranded, and BrandingCategory)

These variables together reveal whether a wholesaler functions as a **brand owner**, **private-label producer**, or **unbranded raw-material supplier**.

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

# Load libraries
library(tidyverse)
library(DBI)
library(stringr)
library(ggplot2)
library(kableExtra)

source("../R_scripts/connect_duckdb_readonly.R")

# product_diversity <- dbGetQuery(con, "
# SELECT
#   LicenseeId,
#   COUNT(DISTINCT ProductId) AS TotalSKUs,
#   COUNT(DISTINCT InventoryType) AS DistinctTypes,
#   SUM(CASE 
#         WHEN InventoryType IN (
#           'Usable Cannabis',
#           'Flower Lot',
#           'Flower Unlotted',
#           'Wet Flower'
#         ) THEN 1 ELSE 0 
#       END) AS FlowerSKUs
# FROM Product_all
# WHERE IsDeleted IS NULL OR IsDeleted NOT IN ('true', 'TRUE')
# GROUP BY LicenseeId
# ")

# Load product diversity with first/last created and sold dates
product_diversity <- dbGetQuery(con, "
SELECT
  p.LicenseeId,
  COUNT(DISTINCT p.ProductId) AS TotalSKUs,
  COUNT(DISTINCT p.InventoryType) AS DistinctTypes,
  SUM(CASE 
        WHEN p.InventoryType IN (
          'Usable Cannabis',
          'Flower Lot',
          'Flower Unlotted',
          'Wet Flower'
        ) THEN 1 ELSE 0 
      END) AS FlowerSKUs,
  MIN(p.CreatedDate) AS FirstProductCreated,
  MAX(p.CreatedDate) AS LastProductCreated,
  MIN(s.CreatedDate) AS FirstSaleDate,
  MAX(s.CreatedDate) AS LastSaleDate
FROM Product_clean p
LEFT JOIN Inventory_all i
  ON p.ExternalIdentifier = i.ExternalIdentifier
LEFT JOIN SalesDetail_all s
  ON i.InventoryId = s.InventoryId
WHERE p.IsDeleted IS NULL OR p.IsDeleted NOT IN ('true', 'TRUE')
GROUP BY p.LicenseeId"
)


product_diversity <- product_diversity %>%
  mutate(
    OnlyFlower = (DistinctTypes == 1 & FlowerSKUs == TotalSKUs),
    BroadSKU = (DistinctTypes > 1)
  )

wholesaler_profiles <- product_diversity %>%
  left_join(
    dbGetQuery(con, "
      SELECT LicenseeId, DBA
      FROM Licensee_all
      WHERE IsDeleted IS NULL OR IsDeleted NOT IN ('true', 'TRUE')
    "),
    by = "LicenseeId"
  )

wholesaler_profiles <- wholesaler_profiles %>% distinct()

product_info <- dbGetQuery(con, "
  SELECT 
    LicenseeId,
    ProductId,
    InventoryType,
    Name AS ProductName,
    Description
  FROM Product_all
  WHERE IsDeleted IS NULL OR IsDeleted NOT IN ('true', 'TRUE')
")

# Join licensee DBA
product_info <- product_info %>%
  left_join(wholesaler_profiles %>% select(LicenseeId, DBA), by = "LicenseeId", relationship = "many-to-many") %>%
  mutate(
    self_branded = str_detect(tolower(ProductName), tolower(DBA)) |
                   str_detect(tolower(Description), tolower(DBA))
  )

#Summarize: % of SKUs that are self-branded
self_brand_summary <- product_info %>%
  distinct(LicenseeId, ProductId, self_branded) %>%   # <-- IMPORTANT FIX
  group_by(LicenseeId) %>%
  summarise(
    TotalSKUs = n(),
    SelfBrandedSKUs = sum(self_branded, na.rm = TRUE),
    PercentSelfBranded = round(100 * SelfBrandedSKUs / TotalSKUs, 1),
    .groups = "drop"
  )

#Combine
wholesaler_summary <- wholesaler_profiles %>%
  left_join(self_brand_summary, by = "LicenseeId") %>%
  mutate(
    BrandingCategory = case_when(
      OnlyFlower ~ "Flower Only",
      BroadSKU ~ "Broad SKU Mix",
      TRUE ~ "Other"
    )
  )

wholesaler_summary <- wholesaler_summary %>%
  rename(
    TotalSKUs_Active = TotalSKUs.x,
    TotalSKUs_WithBranding = TotalSKUs.y
  )

```

## Flower Only Licenses  
**Classify Wholesaler SKU Breadth**
```{r, include=TRUE}

product_diversity <- product_diversity %>%
  mutate(
    OnlyFlower = (DistinctTypes == 1 & FlowerSKUs == TotalSKUs),
    BroadSKU = (DistinctTypes > 1)
  )

kable(product_diversity %>% filter(OnlyFlower == 'TRUE'))

```

<br>
These 29 operations, often new or transitional licensees, are monoculture wholesalers — listing only unbranded flower.

They rarely report sales activity, suggesting either limited market access or supply-side testing before formal retail engagement.

In short: they grow, not brand.

<br>
**Market Interpretation**

1. Branding is Concentrated at Retail:  
Retailers and vertically integrated brands hold the consumer mindshare — wholesalers function as the silent partners.

2. The Long Tail of Micro-Wholesalers:  
Dozens of licensees operate with ≤10 SKUs, often supplying regional stores or testing new formulations.

3. Commoditization Continues:  
With “Only Flower” and “Other Material” products still entering the system in 2025, bulk categories remain accessible and price-driven.

4. Innovation Moves Upstream:  
The ‘Other’ wholesalers — clones, concentrates, edibles, genetics — may represent the next phase of differentiation, focusing on value-added inputs rather than finished goods.

## Detect Self Branding SKUs by Category
```{r, include=FALSE}


ggplot(wholesaler_summary, aes(x = BrandingCategory, y = PercentSelfBranded)) +
  geom_boxplot(outlier.shape = NA, fill = "grey90", width = 0.4) +
  geom_jitter(width = 0.25, alpha = 0.4, color = "darkblue") +
  labs(
    title = "Self-Branded SKU % by Wholesaler Type",
    x = "Wholesaler Type",
    y = "% Self-Branded SKUs"
  ) +
  theme_minimal(base_size = 14)


```
**The Broad SKU Majority**  
Roughly 90% of active wholesalers fall into the “Broad SKU Mix” category — distributors with large, multi-type portfolios:

| Example Wholesalers      | Total SKUs | Distinct Types | Percent Self-Branded | Branding Category |
| :----------------------- | ---------: | -------------: | -------------------: | :---------------- |
| *SALISH COAST COMPANY*   |      6,121 |              9 |                  0 % | Broad SKU Mix     |
| *CANNABIS GLASS*         |     26,385 |             11 |                  0 % | Broad SKU Mix     |
| *KALEAFA*                |      5,628 |             14 |                  0 % | Broad SKU Mix     |
| *CANNABIS COUNTRY STORE* |     12,465 |              9 |                  0 % | Broad SKU Mix     |

These firms operate as regional distributors rather than house brands — high SKU diversity, but low self-branding.

They source and resell across categories, serving as aggregators of volume rather than architects of identity.
```{r, include=FALSE}
## Wholesaler Summary
# Explore 'Other' category in more detail
# other_wholesalers <- wholesaler_summary %>%
#   filter(BrandingCategory == "Other") %>%
#   left_join(product_info, select(-DBA), by = "LicenseeId", relationship = "many-to-many") %>%
#   filter(!is.na(InventoryType))  # In case we bring that in later
# 
# # Count product types within 'Other' group
# type_counts <- dbGetQuery(con, "
#   SELECT LicenseeId, InventoryType, COUNT(*) AS TypeSKUCount
#   FROM Product_clean
#   WHERE (IsDeleted IS NULL OR IsDeleted NOT IN ('true', 'TRUE'))
#   GROUP BY LicenseeId, InventoryType
# ")
# 
# # Join to other wholesalers
# other_type_summary <- other_wholesalers %>%
#   select(LicenseeId, DBA = DBA.x) %>%  # rename explicitly
#   distinct() %>%
#   left_join(type_counts, by = "LicenseeId")
# 
# # Table: top product types among 'Other' licenses
# top_types <- other_type_summary %>%
#   group_by(InventoryType) %>%
#   summarise(
#     TotalSKUs = sum(TypeSKUCount, na.rm = TRUE),
#     LicenseeCount = n_distinct(LicenseeId),
#     .groups = "drop"
#   ) %>%
#   arrange(desc(TotalSKUs))

# Show top 10 InventoryTypes
#print(knitr::kable(head(top_types, 10), caption = "Top Inventory Types in 'Other' Wholesalers"))

# Optional: view specific licensees in "Other"
# head(other_wholesalers %>%
#        select(LicenseeId, DBA.x, ProductName, Description, self_branded), 10)
```

```{r, include=FALSE}

# Product type mix among 'Other'
# ggplot(other_type_summary, aes(x = reorder(InventoryType, -TypeSKUCount), y = TypeSKUCount)) +
#   geom_col(fill = "#0072B2") +
#   coord_flip() +
#   labs(
#     title = "Top Product Types Among 'Other' Wholesalers",
#     x = "Inventory Type",
#     y = "SKU Count"
#   ) +
#   theme_minimal(base_size = 13)

```
**The “Other” Category: Niche Wholesalers**
The “Other” branding category, though small, is where Washington’s wholesale diversity shines.
By joining licensees with product inventories and type counts, we identified 11 “Other” wholesalers spanning multiple niche categories.

| Inventory Type             | Total SKUs | Licensee Count |
| :------------------------- | ---------: | -------------: |
| Concentrate for Inhalation |        141 |              2 |
| Plant                      |         48 |             11 |
| Seed                       |         31 |              2 |
| Clones                     |         28 |              6 |
| Other Material Unlotted    |         26 |              6 |
| Liquid Edible              |         24 |              1 |
| Solid Edible               |         19 |              1 |
| Other Material Lot         |          9 |              1 |
| Flower Unlotted            |          7 |              1 |
| Cannabis Mix Packaged      |          3 |              1 |

<br>
Even within this small cohort, there’s remarkable variety.
Some — like CHIEF N CANNABIS LLC — produce wet flower and concentrates, while others (e.g., WASHINGTON CANNABIS CORPORATION, COUNTRY LANE) specialize in plants, clones, and seed material.

These “Other” wholesalers serve critical ecosystem functions: supplying genetics, starter material, and specialty extracts — the inputs that keep both retail and manufacturing tiers viable.

<br>
**Branding Insights**  
A few patterns stand out among all branding categories:

**Self-Branding Remains Rare**:  
Fewer than 3% of wholesalers actively sell under their own label (PercentSelfBranded > 10).
The dominant strategy is contract manufacturing and third-party distribution.

**Vertical Integration Is Selective**:  
A handful of licensees — like KEEP IT GREEN PRODUCTIONS and SALISH COAST COMPANY — show internal production and distribution, but still manage external brands to fill catalog gaps.

**Product Breadth ≠ Brand Identity**:  
Having more SKUs doesn’t imply stronger branding.
Some of the largest distributors (25k–40k SKUs) remain brand-agnostic, highlighting a wholesale-as-infrastructure business model.

As Washington’s cannabis market matures, the spotlight has shifted from retail storefronts to the wholesale backbone supplying them.

**Between 2024 and mid-2025, CCRS data show the emergence of two distinct wholesaler profiles**:

1. Broad SKU Wholesalers — large distributors with hundreds to thousands of SKUs across multiple product types and brand relationships.

2. “Other” Wholesalers — smaller, often single-category operations specializing in plants, clones, edibles, or unbranded flower.

The analysis provide a rare look into how these entities shape the supply-side brand economy of Washington cannabis.

<br>

